Competing Bandits: The Perils of Exploration under Competition

### 2018-09-06 paper/content/introduction.tex
also edits to intro

Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. Being first in the market gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage"). We investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed. However, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm. \asedit{One take-away is that the data advantage is about not just quantity but quality of data.}

\asedit{We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition. In particular, we find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor. We suggest a more refined performance measure, and use it to explain (and frame) some of the competition outcomes.}

>>>>

Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. Being first in the market gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage"). We investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed. However, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm. One take-away is that the data advantage is not just about data quantity but also data quality.

We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition. In particular, we find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor. We suggest a more refined performance measure, and use it to explain (and frame) some of the competition outcomes.
